\chapter*{Preface}

This lab manual, and the labs described herein, have been developed over many years by many people.  The labs are intended as a companion to a course taught from the draft book \emph{A Mathematical Introduction to Classical Control} by Andrew D.~Lewis, and which has been used for some years as the text for MTHE~332 (formerly MATH~332).

The genesis of these labs were experiments performed in The Cave in Jeffery Hall, beginning in the early 1990's, and developed by Jon Davis and Ron Hirschorn.  These labs used a crude \textsf{Linux} real-time implementation, and were a true adventure for students.  Support for this development was provided by the ``Access to Opportunities'' programme of the Ontario government, the BED Fund from the Queen's Engineering Society, and the Department of Mathematics and Statistics.  The current set of labs,~i.e.,~the labs described in this manual, were developed by Andrew Lewis and Neill Patterson in the early 2000's, still on the \textsf{Linux} platform.  The move to equipment supplied by \href{http://www.quanser.com/}{Quanser Consulting}, a Canadian company developing real-time control systems for education and research, was undertaken by Martin Guay in 2004 with the aid of a grant from the McConnell Foundation.  The first working version of this lab manual was produced by Bernard Chan in the summer of~2004.  This version of the lab manual was used and further developed by the lab TAs, Thomas Norman, Jeffrey Calder, Steven Wu, and Jack Horn, over the next few years.  The alterations were smoothed during the summer of~2012 by Daniel Blair and Zachary Kroeze, and the labs are now being used jointly by the courses MTHE~393 and MTHE 332.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "lab-manual"
%%% End: 
